Making Sequential Programs Deterministic

### 2019-04-06 introduction.tex
intro minor edits

We have created an analysis that detects nondeterminism or verifies its
absence in sequential programs.
Our analysis permits a programmer to specify which parts of their program
are intentionally nondeterministic, and it verifies that the remainder is deterministic.
%
If our analysis issues no warnings, then the program produces the same
output when executed twice on the same inputs.  This guarantee is modulo
the limitations of the
analysis, notably concurrency, implicit control flow, and unanalyzed libraries (see \cref{sec:threats}).
%
Our analysis works at compile time, giving a guarantee over every possible
execution of the program, by contrast to unsound dynamic tools that attempt
to discover when a program has exhibited nondeterministic behavior on
specific runs.  There is no need for a special JVM nor rerunning a program
multiple times.
%
Our analysis handles collections that will contain the same values, but
possibly in a different order, on different runs.
%
Our analysis is precise enough for practical use.  It permits calls to
nondeterministic APIs, and only issues a warning if they are used in ways
that may lead to nondeterministic output observed by a user.  Like any
sound analysis, it can issue false positive warnings.

@@

This paper makes the following contributions:
\begin{enumerate}
  \item We designed a type system, named \ourTypeSystem, for expressing determinism properties.

@@

  \item
  \todo{It's a bit weird that we call this a contribution but it is never
    explicitly discussed in the paper.  I think it's probably OK.}
  In a case study, we annotated several libraries, including the collection
    classes and other parts of the JDK (Java's standard library).  This
    provides a formal, machine-readable specification for the libraries, and
    it demonstrates the expressiveness of our type system.

>>>>

We have created an analysis that detects nondeterminism or verifies its
absence in sequential programs.
Our analysis permits a programmer to specify which parts of their program
are intentionally nondeterministic, and it verifies that the remainder is deterministic.
%
If our analysis issues no warnings, then the program produces the same
output when executed twice on the same inputs.  This guarantee is modulo
the limitations of the
analysis, notably concurrency, implicit control flow, and unanalyzed libraries (see \cref{sec:threats}).
%
Our analysis works at compile time, giving a guarantee over every possible
execution of the program, by contrast to unsound dynamic tools that attempt
to discover when a program has exhibited nondeterministic behavior on
specific runs.  There is no need for a custom runtime system nor rerunning a program
multiple times.
%
Our analysis handles collections that will contain the same values, but
possibly in a different order, on different runs.
%
Our analysis is precise enough for practical use.  It permits calls to
nondeterministic APIs, and only issues a warning if they are used in ways
that may lead to nondeterministic output observed by a user.  Like any
sound analysis, it can issue false positive warnings.

@@

This paper makes the following contributions:
\begin{enumerate}
  \item We designed a type system for expressing determinism properties.

@@

  \item
%  \todo{It's a bit weird that we call this a contribution but it is never
%    explicitly discussed in the paper.  I think it's probably OK.}
  In a case study, we annotated several libraries, including the collection
    classes and other parts of the JDK (Java's standard library).  This
    provides a formal, machine-readable specification for the libraries, and
    it demonstrates the expressiveness of our type system.
